PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2023-04-07 refereeResponse.txt

606: How can a mass be expressed in units of feeding zone size?
	We are solving equation 11 for \tilde{b} and using the initial local surface density, along with the final particle mass to calculate the feeding zone size. We tried to explain this starting on line 611, but it sounds like some further clarification is needed. Some additional text has been added to the paragraph following equation 11 to explain where the value for \tilde{b} comes from.
620: What is the intuition behind tilde b, how exactly is it defined? Since it is a dimensionless quantity, calling it an orbital separation is not intuitive to me. Can tilde b be converted to a physical size? Also, I struggle to interpret the critical tilde be of 2*sqrt(3). Do I understand correctly that a body with tilde b less than 2 sqrt(3) is associated with a feeding zone so small that close encounters are not possible, and thus are not expected to grow further?
	This is buried in some of the references in this section (Nazakawa + Ida 1988, Kokubo 1998, Kokubo 2002), but I'm realizing this is not stated explicitly anywhere in the paper so I will do so: \tilde{b} is a length in units of the local Hill radius. 
	The \tilde{b} = 2\sqrt{3} value is important because this is the smallest possible feeding zone size that a body can have. As eccentricity decreases, the feeding zone size decreases, but has a minimum value of 2\sqrt{3} for (e,i) = (0,0) (See section 6 of Nazakawa + Ida 1988 for an explanation as to why). Therefore, a protoplanet undergoing oligarchic growth (in which gravitational scattering events are frequent and keeps the largest bodies on circular orbits) will not accrete more than 2\sqrt{3}*r_h*Sigma worth of material from its surroundings. The key takeaway from the bottom panel of figure 5 is that there are bodies interior to 60d that must have accreted material from a much wider annulus than this, and so oligarchic growth is not operating here.
	TODO: Update text somewhere?

@@

632: Why are the bodies expected to follow the tilde b = 2sqrt(3) from a theoretical perspective? Also, I'm unconvinced that the bodies do that at all in the presented simulations - they fall below this value but don't follow it.
	See response to 620. You are right that the outer disk bodies do not all reach \tilde{b} = 2\sqrt{3}. TODO: Should probably mention that this should be interpreted as a maximum. Not feasible to run the outer disk to completetion (especially in the fdSteep case).

@@

640: If anything the opposite seems to be true, no? Maybe have a specific value for the expected eccentricity of protoplanets with and without inelastic damping.
	TODO
644: If I look at the contour map by eye, the eccentricity distribution doesn't seem all that different in the inner vs outer disk. Maybe the difference can be quantified?
	You are right that the variation in eccentricity is not huge. For the model used by our initial condition generator, the mean eccentricity is a factor of 4x larger at the outer disk edge, compared to the inner disk edge. This is now stated in the text.

@@

647: This is an interesting point. I'd be interested in seeing a comparison between inner and outer disk accounting for that difference in time scales. Is it possible to include such a comparison maybe with a figure?
	Yes! See our updated version of figure 6. The time is now shown in units of the local accretion timescale, and it is now apparent that the planetesimal surface density curves all follow the same trajectory exterior of 60d.

@@

Figure 6:
This figure could be improved if each line was colored by the semi-major axis of the radius bin, and if there was a corresponding colorbar. The difference between the 60 day boundary could still be indicated with an additional line.
	See our response to the previous concern, along with our response to major comment #1.
	TODO: Update text to reflect changes to figure 6.
	TODO: Run fdhi for a thousand steps at higher output cadence, so the innermost curve starts sooner

@@

665: I struggle to see that there is a significant qualitative difference in the behavior in the lines just inside and just outside the boundary of 60 orbits. The first dashed line also seems to converge to a sigma/Sigma \sim 5 %, not much below the first solid line.
	See our response to the previous concern, along with our response to major comment #1.

@@

668: add reference. Also footnote 5 can probably go in the main text. Can the authors provide an equation that converts tilde b to units of r_h?
	\tilde{b} is already in units of r_h. This is now clarified near line 620 in the text. The 30 percent value simply comes from removing the size of the feeding zone (2\sqrt{3}) from the spacing between oligarchs (10). This is now clarified in the text here. I think the footnote text can be removed after adding this clarification.

@@

682: This statement seems only to be true for fdshallow. The others have a spread in eccentricity that covers two orders of magnitude...
	I think we need to change the wording here. What we were trying to say is: the resulting a-e distributions for the embryos between all of the simulations are quite similar (with the exception of fdLo), despite the fact that the a-e distributions of the initial conditions are different.

@@

685: It is not clear to me why more inelastic collisions lead to a more sharply peaked eccentricity distribution, and not just more damping like the authors argue for the fdLo simulation.
	I'm not sure what you mean by 'more shaply peaked'? The argument we're trying to make here is this: In the case that the initial surface density is large (e.g. the inner part of the fdSteep simulation), the initial eccentricity dispersion is larger (because the viscous stirring timescale is shorter and so we start out with a larger <e^2>^1/2 to make t_vs = t_gas). The increased surface density leads to more collisions, which leads to more inelastic damping. So the larger eccentricity dispersion that we started with quickly damps out.

@@

701: Again, I don't see that the value for tilde b approaches 2sqrt(3) beyond P=60d orbits. The bodies have smaller feeding zones than tilde b = 2sqrt(3)
	Perhaps if we ran some of these simulations for longer, the embryos beyond 60d would all reach \tilde{b} = 2\sqrt{3}. But this requires an intractible number of timesteps, particularly for the fdSteep simulation, where the surface density in the outer disk is quite low. I think the best we can say here is that none of the embryos exceed \tilde{b} = 2\sqrt{3} which is consistent with the oligarchic growth model.

@@

702: Are they that similar? Maybe this statement can be more quantitative? To me it looks like the values fall below 2sqrt(3) at 40 days for fdHiShallow and at 70 days for fdLo. The interpretation that the boundary is only driven by the local value for alpha seems somewhat speculative. I'd also be interested how randomness affects this, i.e. if the simulation is initialized with a different random seed, by how much does the final distribution change?
	We were saving these for the follow up paper, but we do have multiple versions of fdHi, fdSteep and fdShallow! These have now been added to figure 8. The added data does make it clear that the boundary is quite fuzzy. I've also plotted the mean and standard deviation of the \tilde{b} in a number of orbital period bins. I think this makes the accretion boundary a little more apparent. With the exception of maybe the fdShallow simulation, the locations in the disk where \tilde{b} drops below 2\sqrt{3} are broadly consistent, which supports our claim that the boundary location is driven by the local value of alpha.

### 2023-04-07 main.tex
Changes through section 5.4

We show the final state of the ``fdHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel,
the initial (contours) and final (points) state of the simulation is
shown in the orbital period-eccentricity plane. The size of the points
indicates the relative mass of the bodies. In the bottom panel, the
final masses of the bodies (in units of feeding zone size) are shown
as a function of orbital period. The feeding zone size $\tilde{b}$ is
calculated according to \citet{kokubo02}
\begin{equation}\label{eq:iso}
	M = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $M$ is taken to be the final mass of each body. \textbf{The y-values in the bottom panel of figure \ref{fig:fulldisk_e_m} are calculated by solving equation \ref{eq:iso} for $\tilde{b}$, and inputting the initial surface density and final particle mass into the expression.} Here, we assume that each embryo has reached its isolation mass, and 
that differences in mass are driven by differences in the initial surface density at the embryo's current location, along with the 
size of its feeding zone, which is set by the dynamics relevant at that location in the disk. By plotting the derived value of $
\tilde{b}$ as a function of orbital period, differences in the dynamical interactions at different locations of the disk are made more 
clearly visible. For two bodies on circular, non-inclined orbits,  $\tilde{b} = 2
\sqrt{3}$ (indicated by the horizontal dashed line) is the smallest orbital separation that produces a non-negative Jacobi energy
and permits a close encounter (see \citet{naka88}). In typical oligarchic growth simulations \citep{kokubo98}, protoplanets tend 
to space themselves apart by $\tilde{b} = 10$, although it should be noted that they do not consume all of the planetesimals 
within this distance.

@@

In section \ref{sec:narrow}, we postulated that the increased importance of inelastic damping in the inner, non-oligarchic growth 
region of the disk should lower the overall eccentricity of the protoplanets there. This behavior is not immediately apparent in the 
top panel of figure \ref{fig:fulldisk_e_m}. There are, however, a couple of factors in the wide disk simulations that could make this 
extra dynamical cooling mechanism difficult to see. Firstly, the initial eccentricity distributions of the inner and outer disk are 
different because of the dependence of the viscous stirring and gas drag timescales on orbital period. Additionally, the 
protoplanet formation timescales for the inner and outer disk are vastly different, making a comparison between these regions at 
the same moment in time somewhat inappropriate.

To ensure that the boundary seen near 60 days in orbital period is not
simply a transient product of the inside-out growth throughout the
disk, we examine the time evolution of $\sigma/\Sigma$, which compares the planetesimal and total solid surface density at 
multiple orbital periods. In figure \ref{fig:pl_frac_time}, the value of
$\sigma/\Sigma$ is plotted as a function of time in 10 orbital period
bins, each with a width of 10 days.
Bins interior to 60 days are shown by a dashed line, while those exterior are shown as a solid line. In all radial bins, the 
planetesimal surface density is approaching an asymptotic value. In the inner disk, this value asymptotes to zero as the 
planetesimal population entirely depletes. In the outer disk, dynamical friction between the embryos and planetesimals 
eventually throttles subsequent accretion and leaves $\sim$ 10 percent or more of the mass surface density as planetesimals. It 
should be noted that in a typical oligarchic growth scenario, where protoplanets space themselves apart by 10 $r_{h}$ and settle 
onto circular orbits (giving $\tilde{b} = 2 \sqrt{3}$), roughly 30 percent of the planetesimals should remain out of reach of the 
protoplanets\footnote{A feeding zone width of $ \tilde{b} = 2 \sqrt{3}$ covers roughly 70 percent of the 10 $r_{h}$ space between the protoplanets.}.

Next, we investigate how the resulting planetesimal and protoplanet distribution
changes as we vary the initial solid surface density profile.
The final orbital period-eccentricity state of the
particles in the fdHi, fdHiSteep, fdHiShallow and fdLo simulations are shown in figure \ref{fig:surfden_profiles}, with point sizes 
indicating the relative masses of the bodies. In all cases, the inner disk is largely depleted of planetesimals, while the outer disk 
contains a bimodal population of planetesimals and embryos, with a clear separation in eccentricity between the two. Despite 
having significantly different masses, the planetary embryos formed in all simulations have remarkably similar eccentricities. This 
is likely due to the fact that inelastic collisions play a more significant role where the solid surface density is highest, which 
offsets the fact that the initial bodies started off in a dynamically hotter state (due to the increased effectiveness of viscous stirring). The only exception to this is the fdLo simulation, 
where the resulting eccentricities are a couple orders of magnitude smaller. Inelastic damping likely plays an even more 
significant role here, due to the much larger masses of the initial planetesimals.

>>>>

We show the final state of the ``fdHi'' simulation in figure \ref{fig:fulldisk_e_m}. In the top panel,
the initial (contours) and final (points) state of the simulation is
shown in the orbital period-eccentricity plane. The size of the points
indicates the relative mass of the bodies. In the bottom panel, the
final masses of the bodies (in units of feeding zone size) are shown
as a function of orbital period. The feeding zone size $\tilde{b}$ \textbf{is described in units of the local Hill radius and is} calculated according to \citet{kokubo02}
\begin{equation}\label{eq:iso}
	M = \left[ \frac{\left( 2 \pi a^2 \Sigma \tilde{b} \right)^3}{3 M_{*}} \right]^{1/2},
\end{equation}
where $M$ is taken to be the final mass of each body. \textbf{The y-values in the bottom panel of figure \ref{fig:fulldisk_e_m} are calculated by solving equation \ref{eq:iso} for $\tilde{b}$, and inputting the initial surface density and final particle mass into the expression.} Here, we assume that each embryo has reached its isolation mass, and 
that differences in mass are driven by differences in the initial surface density at the embryo's current location, along with the 
size of its feeding zone, which is set by the dynamics relevant at that location in the disk. By plotting the derived value of $
\tilde{b}$ as a function of orbital period, differences in the dynamical interactions at different locations of the disk are made more 
clearly visible. For two bodies on circular, non-inclined orbits,  $\tilde{b} = 2
\sqrt{3}$ (indicated by the horizontal dashed line) is the smallest orbital separation that produces a non-negative Jacobi energy
and permits a close encounter (see \citet{naka88}). In typical oligarchic growth simulations \citep{kokubo98}, protoplanets tend 
to space themselves apart by $\tilde{b} = 10$, although it should be noted that they do not consume all of the planetesimals 
within this distance.

@@

In section \ref{sec:narrow}, we postulated that the increased importance of inelastic damping in the inner, non-oligarchic growth 
region of the disk should lower the overall eccentricity of the protoplanets there. This behavior is not immediately apparent in the 
top panel of figure \ref{fig:fulldisk_e_m}. There are, however, a couple of factors in the wide disk simulations that could make this 
extra dynamical cooling mechanism difficult to see. Firstly, the initial eccentricity distributions of the inner and outer disk are 
different because of the dependence of the viscous stirring and gas drag timescales on orbital period. \textbf{The mean eccentricity at the outer disk edge is 4x larger than at the inner disk edge.} Additionally, the 
protoplanet formation timescales for the inner and outer disk are vastly different, making a comparison between these regions at 
the same moment in time somewhat inappropriate.

To ensure that the boundary seen near 60 days in orbital period is not
simply a transient product of the inside-out growth throughout the
disk, we examine the time evolution of $\sigma/\Sigma$, which compares the planetesimal and total solid surface density at 
multiple orbital periods. In figure \ref{fig:pl_frac_time}, the value of
$\sigma/\Sigma$ is plotted as a function of time in 10 orbital period
bins, each with a width of 10 days.
Bins interior to 60 days are shown by a dashed line, while those exterior are shown as a solid line. In all radial bins, the 
planetesimal surface density is approaching an asymptotic value. In the inner disk, this value asymptotes to zero as the 
planetesimal population entirely depletes. In the outer disk, dynamical friction between the embryos and planetesimals 
eventually throttles subsequent accretion and leaves $\sim$ 10 percent or more of the mass surface density as planetesimals. It 
should be noted that in a typical oligarchic growth scenario, where protoplanets space themselves apart by 10 $r_{h}$ and settle 
onto circular orbits (giving $\tilde{b} = 2 \sqrt{3}$), roughly 30 percent \textbf{($2\sqrt{3}/10 \simeq 0.3$)} of the planetesimals
should remain out of reach of the protoplanets.

Next, we investigate how the resulting planetesimal and protoplanet distribution
changes as we vary the initial solid surface density profile.
The final orbital period-eccentricity state of the
particles in the fdHi, fdHiSteep, fdHiShallow and fdLo simulations are shown in figure \ref{fig:surfden_profiles}, with point sizes 
indicating the relative masses of the bodies. In all cases, the inner disk is largely depleted of planetesimals, while the outer disk 
contains a bimodal population of planetesimals and embryos, with a clear separation in eccentricity between the two. Despite 
having significantly different masses, the \textbf{semimajor axis-eccentricity} distributions of the planetary embryos formed in all simulations \textbf{are} remarkably similar. This 
is likely due to the fact that inelastic collisions play a more significant role where the solid surface density is highest, which 
offsets the fact that the initial bodies started off in a dynamically hotter state (due to the increased effectiveness of viscous stirring). The only exception to this is the fdLo simulation, 
where the resulting eccentricities are a couple orders of magnitude smaller. Inelastic damping likely plays an even more 
significant role here, due to the much larger masses of the initial planetesimals.
